Depression and suicidality as costly signals of need

### 2021-04-02 Paper.rmd
Worked on signaling section - Added signal vs. cue paragraph - Minor edits combining both bargaining models

@@

One possible reason for the connection between depression and social adversity is that depression functions to elicit increased support from others in times when an individual cannot solve substantial fitness-relevant challenges on their own. The bargaining model of depression is based on the idea that for much of our evolutionary history the small groups humans lived in created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop (Hagen, 2003). In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others value the depressed person's cooperation sufficiently.

One way depression and suicidal behavior are thought to lead to beneficial treatment is by reliably demonstrating that one’s need is genuine by making one’s private information about their condition public. This information is necessary due to the ease in which signals can be faked (Zahavi, 1975) and the costs of accepting a deceptive signal as reliable (cite), both of which have favored adaptations to spot fake signals in many species (Hasson, 1994) and demand trustworthy information before believing a given signal (check Hasson and cite + try to find others).

Although there may be other ways for honest signals evolve (cite), one of the most widely accepted (is this fair?) means is through costly signaling. The evolution of costly signals is possible whenever the signal in question is comparatively cheaper to produce for those who have the underlying quality compared to those who lack it (Grafen 1990). In these instances, individuals with the quality maximize their fitness at greater levels of investment in the signal compared to those who lack the quality or possess it to a lesser degree (Gangestad and Scheyd 2005; Zahavi 1977). The end result of this is expected to be a signal that either tends to not be worth faking (Gangestad and Scheyd 2005) or is impossible to fake to a large degree due to selection against fakers constraining the physiology behind the signal (Zahavi 1977).

@@

>>>>

- emphasize conflict more

@@

- add suicide as a pure signal
- merge the hypotheses more

One possible reason for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others in times when an individual cannot solve substantial fitness-relevant challenges on their own (cite). Referred to as the bargaining models of depression and suicide, the idea is that the small groups humans lived in for much of our evolutionary history created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop (Hagen, 2003). In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others valued an individual's cooperation sufficiently.

One way this approach expects depression and suicidal behavior to lead to beneficial treatment is their ability to reliably demonstrate that one’s need is genuine by making one’s private information about their condition public (cite). This information is necessary due to the ease in which signals can be faked (Zahavi, 1975) and the costs of accepting a deceptive signal as reliable (cite), both of which have favored adaptations to spot fake signals in many species (Hasson, 1994) and demand trustworthy information before believing a given signal (check Hasson and cite + try to find others).

Although there may be other ways for honest signals evolve (cite), one of the most widely accepted (is this fair?) means is through costly signaling. The evolution of costly signals is possible whenever the signal in question is comparatively cheaper to produce for those who have the underlying quality being compared to those who lack it (Grafen 1990). In these instances, individuals with the quality maximize their fitness at greater levels of investment in the signal compared to those who lack the quality or possess it to a lesser degree (New paper from Barclay paper; Zahavi 1977). The end result of this is expected to be a signal that either tends to not be worth faking (New paper from Barclay paper) or is impossible to fake to a large degree due to selection against fakers constraining the physiology behind the signal (Zahavi 1977).

@@

As with any study examining signaling, distinguishing between the information provided by signals vs. cues is a major challenge. A signal differs from a cue in that it is intentionally sent by an organism to alter the behavior or thoughts of other individuals, while a cue is any piece of information an organism uses to alter its behavior or physiological state that was created without the intent to do so (Maynard Smith & Harper, 1995). When it comes to this study, suicide attempts are more likely to take the form of a pure, external signal than depression symptoms. This is because behavioral shifts motivated by low mood for non-signaling benefits would convey information as cues in addition to what might exist for signaling purposes. In other words, the signaling portion of depression may come from greater investment in symptoms than would be optimal for motivating behavior changes alone instead of the creation of entirely new symptoms. As most suicide attempts likely do not have non-signaling functions (see DC for possible exceptions), it is likely much easier to interpret the effect of signaling compared to depression.
